---
title: "DATA607_Project1"
author: "Harpreet Shoker"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### PROJECT 1
In this project, we’re given a text file with chess tournament results where the information has some structure.Here is how the structure looks like in the below table
```{r }
# Load packages
library(stringr)
library(data.table)
# Reading the data file and skipping header block
chessTournament <- read.csv("https://raw.githubusercontent.com/Harpreet1984/DATA607/master/tournamentinfo.txt", sep = ",", skip = 3)

# Sample raw input data
head(chessTournament)
```

#Remove the dashed lines
```{r }
chessTournamentFiltered <- chessTournament[!str_detect (chessTournament[,1], "^-+$"),]
```

#combining multiple customer rows into one
```{r }
customerCombined <- str_c(chessTournamentFiltered[seq(1, length(chessTournamentFiltered), 2)], chessTournamentFiltered[seq(2, length(chessTournamentFiltered), 2)], sep = "")
customerCombined <- data.frame(str_split_fixed(customerCombined, "\\||/|->", n = Inf))
```

#update column headers, setup computational columns
```{r}
colnames(customerCombined) <- c("PlayerId",
 "PlayerName",           
 "PlayerPoints",         
 "Game1",                 
 "Game2",
 "Game3",
 "Game4",
 "Game5",
 "Game6",
 "Game7",
 "PlayerState",          
 "UscfId" ,
 "PlayerRatingPre",
 "PlayerRatingPost",
 "nValue",
 "value1",
 "value2",
 "value3",
 "value4",
 "value5",
 "value6",
 "value7"
)
 
customerCombined$PlayerRatingPre <- as.numeric(str_replace(customerCombined$PlayerRatingPre, "R: *([:digit:]+)(P.*)?", "\\1"))

customerCombined$PlayerRatingPost <- as.numeric(str_replace(customerCombined$PlayerRatingPost, "([:digit:]+)(P.*)?", "\\1"))

customerCombined$`AvePreRating` <- as.numeric(0)
```

#Create derived table, for getting average opponent.
```{r}
opp1 <- data.frame(as.numeric(str_extract_all(customerCombined$Game1,"[[:digit:]]{1,}")))
opp2 <- data.frame(as.numeric(str_extract_all(customerCombined$Game2,"[[:digit:]]{1,}")))
opp3 <- data.frame(as.numeric(str_extract_all(customerCombined$Game3,"[[:digit:]]{1,}")))
opp4 <- data.frame(as.numeric(str_extract_all(customerCombined$Game4,"[[:digit:]]{1,}")))
opp5 <- data.frame(as.numeric(str_extract_all(customerCombined$Game5,"[[:digit:]]{1,}")))
opp6 <- data.frame(as.numeric(str_extract_all(customerCombined$Game6,"[[:digit:]]{1,}")))
opp7 <- data.frame(as.numeric(str_extract_all(customerCombined$Game7,"[[:digit:]]{1,}")))

opps <- cbind(customerCombined$PlayerId, customerCombined$PlayerName, opp1, opp2, opp3, opp4, opp5, opp6, opp7)
names(opps) <- c("Player Id","Player Name","Opp 1","Opp 2","Opp 3","Opp 4","Opp 5","Opp 6","Opp 7")

opps[is.na(opps)] <- as.numeric(0)

```

#Extract opponent average points
```{r}
for (k in 3:9){
  for (j in 1:dim(customerCombined)[1]){
    for (i in 1:dim(customerCombined)[1]){
      if (as.numeric(opps[j,k]) == as.numeric(customerCombined$PlayerId[i])){
         customerCombined$AvePreRating[j] <- as.numeric(customerCombined$AvePreRating[j]) + as.numeric(customerCombined$PlayerRatingPre[i])
      }
    }  
  }
}  

customerCombined$AvePreRating  <- round(as.numeric(customerCombined$AvePreRating) / 7,0)

head(customerCombined)
```

Our aim here is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the player
1. Player Name
2. Player State
3. Total points
4. Player Pre-Rating,
5. Average Opponents Raing

#generate csv file
```{r }
write.table(customerCombined[, c(2, 3, 11, 13, 24)], file = "Harpreet-chesstournamentInfo.csv",row.names=FALSE, na="",col.names = c("Player Name", "Points", "Player State", "Pre Rating", "Opponent Avg Rating"), sep=",")
```

