---
title: "DATA607_Assignment3"
author: "Harpreet Shoker"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

### Environment Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(stringr)
```

### Extracting data from example from book
```{r }
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555
-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,
Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

#### (a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name

```{r }
names_new <- str_replace(name, pattern = "Rev. |Dr. |Mr. |Mrs. |Ms ", replacement = ""); #replacing the titles 
firstname = sapply(strsplit(names_new, ' '), function(x) x[1])
lastname = sapply(strsplit(names_new, ' '), function(x) x[length(x)])
full_names <- cbind(firstname,lastname)
full_names
```

#### (b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r }
title <- str_detect(name, "^.{3}\\.|^.{2}\\.")
check_title <- cbind(name, title = title)
check_title
```

#### (c) Construct a logical vector indicating whether a character has a second name.

```{r }
second_name <- ifelse(str_detect(name,"[A-Z]\\.") == TRUE, "YES", "NO")
mid_name <- cbind(name,second_name)
mid_name
```

#### 4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

#### (a) [0-9]+\\$

This expression matches any digit or digits  between 0 to 9 ending with $ sign
here is the example
```{r }
data1 <- c("Example for the above is expression is 0$","46758$","65472999")
unlist(str_extract_all(data1, "[0-9]+\\$"))

```

#### (b) \\b[a-z]{1,4}\\b

This expression matches 1 to 4 character lower case words in input string
```{r }
data2 <- c("Example for the above is expression is one four","1","Two","one")
unlist(str_extract_all(data2, "\\b[a-z]{1,4}\\b"))

```


#### (c) .*?\\.txt$

This expression matches any string ending with .txt including the string that equals .txt. Can be used to find text files in a file listing.

```{r }
data2 <- c("Example for the above is expression is one four.txt","1","Two","one.txt")
unlist(str_extract_all(data2, ".*?\\.txt$"))

```

#### (d) \\d{2}/\\d{2}/\\d{4}

This expression matches 2 digits followed by a / followed by another 2 digits followed by a / and finally followed by four digits
We can say date format

```{r }
data2 <- c("18/07/1984","1","Two","one.txt")
unlist(str_extract_all(data2, "\\d{2}/\\d{2}/\\d{4}"))
```

#### (e) <(.+?)>.+?</\\1>

This expression matches a pair of openning and closing tags - one or more characters enclosed with < > brackets /tags .

```{r }
data2 <- c("18/07/1984","1","Two","one.txt","<a>Harpreet<a>test</a></a>")
unlist(str_extract_all(data2, "<(.+?)>.+?</\\1>"))
```


#### Extra credit Question --- 9. The following code hides a secret message. Crack it with R and regular expressions.Hint: Some of the characters are more revealing than others! The code snippet is alsoavailable in the materials at www.r-datacollection.com.clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr


```{r}
secret_message <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
paste(unlist(str_extract_all(secret_message, "[:upper:]|[:punct:]")))
```
      